---
title: "Reproducable Research: Course Project 1"
author: "Mattias Östlund"
date: "12 juni 2016"
output: 
        html_document:
                toc: true
                toc_depth: 2
                theme: cosmo
                highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation

Packages used in this document are:

*   downloader
*   tidyr
*   dplyr
*   lubridate
*   ggplot2
```{r echo = FALSE, message = FALSE}
library(downloader)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
```


Download, unzip and load the data to d
```{r eval = TRUE }
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download(url, dest="dataset.zip", mode="wb") 
unzip ("dataset.zip", exdir = "./")
d <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
### Prepare dataset

```{r }
dd <- d %>% 
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarise(no_of_steps = sum(steps))
```


### Result

Histogram
```{r }
hist(dd$no_of_steps, main = "Total Steps", xlab = "No of Steps", col = "blue")
```

The average of no of steps per day is **`r as.integer(mean(dd$no_of_steps))`** and meadian is **`r median(dd$no_of_steps)`**



## Daily activity pattern
### Prepare dataset

```{r }
dd <- d %>% 
        filter(!is.na(steps)) %>%
        group_by(interval) %>%
        summarise(no_of_steps = mean(steps))

max_interval <- dd[which.max(dd$no_of_steps),]
average_steps <- as.integer(mean(d$steps, na.rm=TRUE))
```


### Result

```{r }
plot(dd$interval, dd$no_of_steps, type = "l", main = "Average No. of Steps per Interval", xlab = "Interval", ylab ="No. of Steps", col = "blue")
```

The interval with max steps is **`r max_interval$interval`** and has **`r as.integer(max_interval$no_of_steps)`** steps.



## Imputing missing values


Total numbers of rows of missing data (NA) are **`r sum(is.na(d$steps))`**.

Strategy to replace NA are taking the average steps for whole data set. Use the variable, average_steps which has the value of **`r average_steps`**.

### Replace NA's

Create a new dataset d_na whwew NA's are replaced with **`r average_steps`**.
```{r}
d_na <- d
d_na$steps[which(is.na(d_na$steps))] <- average_steps
```

Histogram with replace NA's
```{r }
dd_na <- d_na %>% 
        group_by(date) %>%
        summarise(no_of_steps = sum(steps))
hist(dd_na$no_of_steps, main = "Total Steps", xlab = "No of Steps", col = "blue")
```

The average of no of steps per day is **`r as.integer(mean(dd_na$no_of_steps))`** and meadian is **`r median(dd_na$no_of_steps)`**

## Differences in activity patterns between weekdays and weekends

```{r }
dd_na <- d_na %>%
  mutate(wd = weekdays(ymd(as.character(date)),abbreviate = TRUE), wtype = factor(ifelse(wd %in% c("Mån", "Tis", "Ons", "Tor", "Fre"), "weekday" , "weekend"))) %>%
  group_by(wtype, interval) %>%
  summarise(no_of_steps = mean(steps))

g <- ggplot(data=dd_na, aes(x=interval, y=no_of_steps)) 
g <- g +  geom_line(color="blue")
g <- g + facet_wrap( ~ wtype, ncol = 1)
g <- g + ggtitle("Activity patterns between weekdays and weekends")
g <- g + xlab("Interval")
g <- g + ylab("No of Steps")
g
```

### End of Report